---
title: "Part A - SQL"
execute:
  echo: false
---

```{r Install dependencies}
#| echo: false
#| warning: false
if(!require('tidyverse'))
  packages.install('tidyverse')

if(!require('RSQLite'))
  packages.install('RSQLite')
```

# Core SQL Queries

### 1. List the total budget allocated for projects in each country, along with the count of projects per country. Display sorted by the total budget in descending order

@tbl-projectbudgetspercountry presents the total budget allocated for projects in each country, along with the number of projects per country

```{r}
#| label: tbl-projectbudgetspercountry
#| tbl-cap: Project count and budget per country
#| echo: false
#| warning: false
#| cache: true
library("RSQLite")

## connect to db
con <- dbConnect(drv=RSQLite::SQLite(), dbname="ICA_2023.sqlite")

projectBudgets <- dbGetQuery(
  con,
  'SELECT
      CustomerCountry AS Country,
      SUM(Budget) AS \'TotalBudget\',
      COUNT(ProjectID) AS "ProjectCount"
  FROM Projects
      JOIN Customers C ON
          Projects.CustomerID = C.CustomerID
  GROUP BY CustomerCountry
  ORDER BY TotalBudget DESC;'
)

dbDisconnect(con)

# Show up to 7 rows. If data has more than 7 entries, show the first 8 with pagination
if(nrow(projectBudgets) <= 7) {
  knitr::kable(head(projectBudgets, 7), align='lcc')
} else {
  DT::datatable(projectBudgets, options = list(pageLength = 8, columnDefs = list(list( className = 'dt-center', targets = 2:3))))
}
```

### 2. List the average development time for projects, categorized by the number of assets used

@tbl-devtimebyassetsused presents the average development time for projects, grouped by the number of assets used

```{r}
#| label: tbl-devtimebyassetsused
#| tbl-cap: Average development time for projects, grouped by number of assets used
#| echo: false
#| warning: false
#| cache: true
library("RSQLite")

## connect to db
con <- dbConnect(drv=RSQLite::SQLite(), dbname="ICA_2023.sqlite")

avgDevTimeByAssetCount <- dbGetQuery(
  con,
  'SELECT
      AVG(DaysTaken) AS AverageDevTime,
      AssetCount
  FROM (
      SELECT
          COUNT(AssetID) AS AssetCount,
          (JULIANDAY(EndDate) - JULIANDAY(StartDate)) AS DaysTaken
      FROM Projects
          LEFT JOIN Assets A ON
              Projects.ProjectID = A.ProjectID
      GROUP BY Projects.ProjectID
  )
  GROUP BY AssetCount;'
)

dbDisconnect(con)

# Show up to 7 rows. If data has more than 7 entries, show the first 8 with pagination
if(nrow(avgDevTimeByAssetCount) <= 7) {
  knitr::kable(head(avgDevTimeByAssetCount, 7), align='c')
} else {
  DT::datatable(avgDevTimeByAssetCount, options = list(pageLength = 8, columnDefs = list(list( className = 'dt-center', targets = 0:2))))
}
```

### 3. List the top three developers based on the number of successful projects they've been involved in. Display the results

@tbl-mostsuccessfuldevs presents the 3 developers with the highest number of completed projects

```{r}
#| label: tbl-mostsuccessfuldevs
#| tbl-cap: Developers with the highest number of successful projects
#| echo: false
#| warning: false
#| cache: true
library("RSQLite")

## connect to db
con <- dbConnect(drv=RSQLite::SQLite(), dbname="ICA_2023.sqlite")

mostSuccessfulDevs <- dbGetQuery(
  con,
  'SELECT
      Name,
      COUNT(ProjectDevelopers.ProjectID) AS SuccessfulProjects
  FROM ProjectDevelopers
      LEFT JOIN Projects P ON
          ProjectDevelopers.ProjectID = P.ProjectID
      JOIN Developers D ON
          ProjectDevelopers.DeveloperID = D.DeveloperID
  WHERE
      Status IS \'Completed\'
  GROUP BY D.DeveloperID
  ORDER BY SuccessfulProjects DESC
  LIMIT 3;'
)

dbDisconnect(con)

# Show up to 7 rows. If data has more than 7 entries, show the first 8 with pagination
if(nrow(mostSuccessfulDevs) <= 7) {
  knitr::kable(head(mostSuccessfulDevs, 7), align='lc')
} else {
  DT::datatable(mostSuccessfulDevs, options = list(pageLength = 8, columnDefs = list(list( className = 'dt-center', targets = 2))))
}
```

# General SQL Concepts

### 1. SELECT with **LIKE** and **OR**

@tbl-mostexperiencesdesignersandengineers presents all developers with "designer" or "engineer" in their specialization, ordered by years of experience

```{r}
#| label: tbl-mostexperiencesdesignersandengineers
#| tbl-cap: Years of experience amongst designers and engineers
#| echo: false
#| warning: false
#| cache: true
library("RSQLite")

## connect to db
con <- dbConnect(drv=RSQLite::SQLite(), dbname="ICA_2023.sqlite")

designersAndEngineersExperience <- dbGetQuery(
  con,
  'SELECT
    Name, Specialization, ExperienceYears
  FROM Developers
  WHERE
      LOWER(Specialization) LIKE \'%designer\'
      OR
      LOWER(Specialization) LIKE \'%engineer\'
  ORDER BY ExperienceYears DESC;'
)

dbDisconnect(con)

# Show up to 7 rows. If data has more than 7 entries, show the first 8 with pagination
if(nrow(designersAndEngineersExperience) <= 7) {
  knitr::kable(head(designersAndEngineersExperience, 7), align = 'lcc')
} else {
  DT::datatable(designersAndEngineersExperience, options = list(pageLength = 8, columnDefs = list(list( className = 'dt-center', targets = 2:3))))
}
```

### 2. *SELECT* with **DISTINCT** and **ORDER BY**

@tbl-projectleads presents the names of all project leads

```{r}
#| label: tbl-projectleads
#| tbl-cap: Project leads
#| echo: false
#| warning: false
#| cache: true
library("RSQLite")

## connect to db
con <- dbConnect(drv=RSQLite::SQLite(), dbname="ICA_2023.sqlite")

projectLeads <- dbGetQuery(
  con,
  'SELECT DISTINCT Name
  FROM Developers
      JOIN ProjectDevelopers PD ON Developers.DeveloperID = PD.DeveloperID
      JOIN Projects P ON PD.ProjectID = P.ProjectID
  WHERE
      LOWER(Role) IS \'lead\'
  ORDER BY Name ASC;'
)

dbDisconnect(con)

# Show up to 7 rows. If data has more than 7 entries, show the first 8 with pagination
if(nrow(projectLeads) <= 7) {
  knitr::kable(head(projectLeads, 7), align = 'c')
} else {
  DT::datatable(projectLeads, options = list(pageLength = 8, columnDefs = list(list( className = 'dt-center', targets = 1))))
}
```

### 3. Subquery with **SELECT**

@tbl-multiprojectdevs presents the names and specializations of all developers who have worked on more than 1 project

```{r}
#| label: tbl-multiprojectdevs
#| tbl-cap: Developers with experience on more than one project
#| echo: false
#| warning: false
#| cache: true
library("RSQLite")

## connect to db
con <- dbConnect(drv=RSQLite::SQLite(), dbname="ICA_2023.sqlite")

# The subquery is justified here because the WHERE condition acts on the 
# computed ProjectCount, which is an aggregate call. This isn't allowed in
# SQLite so a subquery is required

multiProjectDevs <- dbGetQuery(
  con,
  'SELECT
      Name,
      Specialization,
      ProjectCount
  FROM (
      SELECT
          Name,
          Specialization,
          COUNT(ProjectID) AS ProjectCount
      FROM ProjectDevelopers
      JOIN Developers D on D.DeveloperID = ProjectDevelopers.DeveloperID
      GROUP BY D.DeveloperID
  )
  WHERE ProjectCount > 1
  ORDER BY ProjectCount DESC;'
)

dbDisconnect(con)

# Show up to 7 rows. If data has more than 7 entries, show the first 8 with pagination
if(nrow(multiProjectDevs) <= 7) {
  knitr::kable(head(multiProjectDevs, 7), align = 'lcc')
} else {
  DT::datatable(multiProjectDevs, options = list(pageLength = 8, columnDefs = list(list( className = 'dt-center', targets = 2:3))))
}
```
